---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE,
  fig.showtext = TRUE
)
library(cts)
rifampicin <- compound(testthat::test_path("data/Rifampicin-Model.json"))
midazolam <- compound(testthat::test_path("data/Midazolam-Model.json"))

showtext::showtext_opts(dpi = 300)
```

# cts

<!-- badges: start -->

<!-- badges: end -->

The `{cts}` (Contraceptives DDI Trial Simulation Platform) package provides a comprehensive R-based framework for designing and simulating drug-drug interactions (DDI) involving contraceptive drugs using physiologically based pharmacokinetic (PBPK) models. This package allows researchers to conduct *in silico* DDI simulations programmatically, enabling more automated and reproducible workflows. This is particularly useful for performing batch simulations, and integrating DDI modeling into larger R-based data analysis pipelines, thus enhancing flexibility and efficiency in model development and simulation.

#### Key Features

-   Import and Explore Compound Models: Retrieve compound models snapshots from the OSP model library, a URL, or a local file in JSON format.
-   Simulate DDIs: Design and run DDI simulations between compounds, specifying dosing protocols, individual characteristics, and other simulation parameters.
-   Analyze Results: In addition to simulation results, this packages provides an immediate and first glance at the data by providing basic plots and tables.
-   Seamless Integration: Supports the export of simulation snapshots compatible with PK-Sim, enabling further configuration, exploration and analysis.


## Installation

You can install the development version of `cts` from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("pak")
pak::pak("esqLABS/cts")
```

## Example

Below are some examples to demonstrate how to use the package. These steps will guide you through importing compound models, exploring their properties, and simulating drug-drug interactions.

### List Available Compound Building Blocks

The `{cts}`package can interface with the Open Systems Pharmacology (OSP) model library to retrieve compound models. The `list_compounds()` function lists all available compounds in the OSP model library.

```{r example, eval = FALSE, attr.output='style="max-height: 100px;"'}
library(cts)

list_compounds()
```
```{r, echo = FALSE, max.print=10, message=FALSE, warning=FALSE}
# List available compounds building blocks
cat(cli::cli_fmt(list_compounds(), collapse = T))
```


This command outputs a list of compound names available in the OSP model library, which you can select to create your DDI simulations.

### Import Snapshots

The core functionality of the package involves importing compound models from snapshot files. You can import a compound model in various ways:


#### From OSP Qualified Building Blocks

You can directly import pre-qualified compounds from the OSP model library:

```{r, eval=FALSE}
rifampicin <- compound("Rifampicin")
midazolam <- compound("Midazolam")
```

#### From URL

For compound models hosted online, use the URL to import the model:

```{r, eval = FALSE}
compound("https://raw.githubusercontent.com/Open-Systems-Pharmacology/Alfentanil-Model/v2.2/Alfentanil-Model.json")
```


#### From Local File

You can also import compound models from local JSON files:

```{r, eval = FALSE}
compound("path/to/Alfentanil-Model.json")
```


### Explore Snapshots

Once you have imported a compound snapshot, you can explore its properties, such as compound properties, formulations, protocols, predefined simulations and all other information contained in the snapshot. For example, you can explore the formulations of Rifampicin using the following command:

```{r}
rifampicin$protocols[[1]]
```

This command outputs the formulations that are defined for the compound and that can be used in simulations.


## Add new building blocks

Only administration protocols can be altered/added/removed from snapshots

```{r}
my_protocol <- create_protocol("my_protocol",
                type = "oral",
                interval = "single",
                dose = 200,
                water_vol_per_body_weight = 3.5)

my_protocol
```

```{r}
add_protocol(rifampicin, my_protocol)
```

```{r}
purrr::map_chr(rifampicin$protocols, "name")
```


## Create, Parameterize and Run a DDI Project

After importing the necessary compounds, you can create a DDI simulation. Here is a basic example of creating a DDI simulation with a victim compound (e.g., Rifampicin) and a perpetrator compound (e.g., Midazolam):

Some additional options can be passed during ddi creation:

```{r, warning=FALSE}
myDDI <- create_ddi(
  victim = rifampicin, 
  midazolam,
  options = list(
    import_simulations = FALSE,
    create_ddi_simulation = TRUE
  )
)
```

All options can be found in the `?create_ddi` help page.  

By default, the `create_ddi` function will generate a generic and ready-to-run ddi simulation.


## Run the DDI simulations

Simulations defined in snapshots can easily be executed and their results retrieved using the `run_ddi()` function:

```{r}
results <- run_ddi(myDDI)
```
This will return a named list of data frames containing the simulation results for each simulation in the DDI project. You can access the results for a specific simulation using its name as a list index:

```{r, eval = FALSE}
head(results$`Generic DDI simulation`)
```
```{r , echo = FALSE}
knitr::kable(head(results$`Generic DDI simulation`))
```


Results can also be exported as csv and pkml files with the `path` and `exportPKML` arguments:

```{r, eval = FALSE}
results <- run_ddi(myDDI, path = "output_directory/", exportPKML = TRUE)
```

## Get PK Parameters

PK parameters can be computed directly from the simulation results using `run_pk_analysis()` function:

```{r}
pk_analysis <- run_pk_analysis(myDDI)
```


As well as results, the PK statistics can be accessed by simulation name:

```{r, eval = FALSE}
head(pk_analysis$`Generic DDI simulation`)
```
```{r, echo=FALSE}
knitr::kable(head(pk_analysis$`Generic DDI simulation`))
```


## Plot ddi simulations

DDI simulations can be generated using the `plot_ddi_results()` function:

```{r} 
plots <- plot_ddi_results(myDDI)
```

Similarly plots can be accessed and visualized by simulation name:

```{r plot-ddi-results, out.width = "85%", fig.align="center", fig.width=6, fig.asp=0.618, dpi = 300}
plots$`Generic DDI simulation`
```

## Export the DDI project as a Snapshot

You can also export the entire DDI simulation as a snapshot file for future use or for importing into other platforms such as PK-Sim:

```{r, eval=FALSE}
# Export the DDI simulation to a JSON snapshot file
export_ddi(myDDI, "path/to/Rifampicin-Midazolam-DDI.json")
```

## Add a new simulation

New simulations can be defined and added to the DDI object. 
Use create_simulation() to define a new simulation, listing its name, the victim compound, the perpetrator compound, and the individual. 
Then, the simulation can be added to the DDI object using the add_simulation() function. 

By default, all compounds interactions and processes will be added to the simulation.

```{r}
my_sim <- create_simulation(
  simulation_name = "My New Simulation",
  victim = "Rifampicin",
  perpetrators = "Midazolam",
  individual = "European (P-gp modified, CYP3A4 36 h)"
)

# add_simulation(myDDI, my_sim, options = list(add_interactions = TRUE, add_processes = FALSE))

```

### Simulations parameterization

Simulations can be parameterized.


#### Add another compound

```{r}
my_sim$add_compound(compound = "Test compound 2", protocol = "New protocol")
```

#### Change administration protocol

```{r}
set_compound_protocol(my_sim, compound = "Rifampicin", protocol = "iv 300 mg (0.5 h)")
```


## Dependencies

The `{cts}` package relies on several components from the Open Systems Pharmacology Suite:

- **R version 4.4.0 or higher**.
- **External R packages**: `{ospsuite}` for interaction with the PK-Sim core and `{rSharp}` for R and C# integration. These packages are installed automatically when you install `{cts}`.

For more details on the installation and dependency requirements for `{ospsuite}`, please visit the `{ospsuiteR}` GitHub page [here](https://github.com/Open-Systems-Pharmacology/OSPSuite-R).

