---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE,
  fig.showtext = TRUE
)
library(cts)
rifampicin <- compound(testthat::test_path("data/Rifampicin-Model.json"))
midazolam <- compound(testthat::test_path("data/Midazolam-Model.json"))

showtext::showtext_opts(dpi = 300)
```

# cts

<!-- badges: start -->

<!-- badges: end -->

The `{cts}` (Contraceptives DDI Trial Simulation Platform) package provides a comprehensive R-based framework for designing and simulating drug-drug interactions (DDI) involving contraceptive drugs using physiologically based pharmacokinetic (PBPK) models. This package allows researchers to conduct *in silico* DDI simulations programmatically, enabling more automated and reproducible workflows. This is particularly useful for performing batch simulations, and integrating DDI modeling into larger R-based data analysis pipelines, thus enhancing flexibility and efficiency in model development and simulation.

#### Key Features

-   Import and Explore Compound Models: Retrieve compound models snapshots from the OSP model library, a URL, or a local file in JSON format.
-   Simulate DDIs: Design and run DDI simulations between compounds, specifying dosing protocols, individual characteristics, and other simulation parameters.
-   Analyze Results: In addition to simulation results, this packages provides an immediate and first glance at the data by providing basic plots and tables.
-   Seamless Integration: Supports the export of simulation snapshots compatible with PK-Sim, enabling further configuration, exploration and analysis.


## Installation

You can install the development version of `cts` from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("pak")
pak::pak("esqLABS/cts")
```

## Get Started

The `{cts}` package offers a streamlined workflow for drug-drug interaction (DDI) simulations. Here's a quick overview of the main steps:

### 1. Import Compound Models

```{r example, eval = FALSE}
library(cts)

# List available compounds in the OSP model library
list_compounds()

# Import compounds from the OSP model library
rifampicin <- compound("Rifampicin")
midazolam <- compound("Midazolam")

# Alternatively, import from URL or local file
# compound("https://raw.githubusercontent.com/Open-Systems-Pharmacology/Alfentanil-Model/v2.2/Alfentanil-Model.json")
# compound("path/to/Alfentanil-Model.json")
```

### 2. Create a DDI Project

```{r, eval=FALSE}
# Create a DDI project with victim and perpetrator compounds
myDDI <- create_ddi(
  victim = rifampicin,
  midazolam,
  options = list(
    create_ddi_simulation = TRUE # Automatically create a generic simulation
  )
)
```

### 3. Run Simulations and Analyze Results

```{r, eval=FALSE}
# Run simulations
results <- run_ddi(myDDI)

# Calculate PK parameters
pk_analysis <- run_pk_analysis(myDDI)

# Generate plots
plots <- plot_ddi_results(myDDI)

# View a specific plot
plots$`Generic DDI simulation`
```

### 4. Export for Further Analysis

```{r, eval=FALSE}
# Export the DDI project to a JSON file for use in PK-Sim
export_ddi(myDDI, "path/to/Rifampicin-Midazolam-DDI.json")
```

## Learn More

For detailed information on using the `{cts}` package, please refer to the vignettes:

```{r, eval=FALSE}
# Browse all vignettes
vignette(package = "cts")

# View a specific vignette
vignette("working-with-compounds", package = "cts")
vignette("create-building-blocks", package = "cts")
vignette("simulations", package = "cts")
vignette("creating-ddi-projects", package = "cts")
```
