---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE,
  fig.showtext = TRUE
)
library(cts)
rifampicin <- compound(testthat::test_path("data/Rifampicin-Model.json"))
midazolam <- compound(testthat::test_path("data/Midazolam-Model.json"))

showtext::showtext_opts(dpi = 300)
```

# cts

<!-- badges: start -->

<!-- badges: end -->

The `{cts}` (Contraceptives DDI Trial Simulation Platform) package provides a comprehensive R-based framework for designing and simulating drug-drug interactions (DDI) involving contraceptive drugs using physiologically based pharmacokinetic (PBPK) models. This package allows researchers to conduct *in silico* DDI simulations programmatically, enabling more automated and reproducible workflows. This is particularly useful for performing batch simulations, and integrating DDI modeling into larger R-based data analysis pipelines, thus enhancing flexibility and efficiency in model development and simulation.

#### Key Features

-   Import and Explore Compound Models: Retrieve compound models snapshots from the OSP model library, a URL, or a local file in JSON format.
-   Simulate DDIs: Design and run DDI simulations between compounds, specifying dosing protocols, individual characteristics, and other simulation parameters.
-   Analyze Results: In addition to simulation results, this packages provides an immediate and first glance at the data by providing basic plots and tables.
-   Seamless Integration: Supports the export of simulation snapshots compatible with PK-Sim, enabling further configuration, exploration and analysis.


## Installation

You can install the development version of `cts` from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("pak")
pak::pak("esqLABS/cts")
```

## Get Started

### Import Compound Models

Below are some examples to demonstrate how to use the package. These steps will guide you through importing compound models, exploring their properties, and simulating drug-drug interactions.

#### List Available Compound Building Blocks

The `{cts}`package can interface with the Open Systems Pharmacology (OSP) model library to retrieve compound models. The `list_compounds()` function lists all available compounds in the OSP model library.

```{r example, eval = FALSE, attr.output='style="max-height: 100px;"'}
library(cts)

list_compounds()
```
```{r, echo = FALSE, max.print=10, message=FALSE, warning=FALSE}
# List available compounds building blocks
cat(cli::cli_fmt(list_compounds(), collapse = T))
```


This command outputs a list of compound names available in the OSP model library, which you can select to create your DDI simulations.

#### Import Snapshots

The core functionality of the package involves importing compound models from snapshot files. You can import a compound model in various ways:


##### From OSP Qualified Building Blocks

You can directly import pre-qualified compounds from the OSP model library:

```{r, eval=FALSE}
rifampicin <- compound("Rifampicin")
midazolam <- compound("Midazolam")
```

##### From URL

For compound models hosted online, use the URL to import the model:

```{r, eval = FALSE}
compound("https://raw.githubusercontent.com/Open-Systems-Pharmacology/Alfentanil-Model/v2.2/Alfentanil-Model.json")
```


##### From Local File

You can also import compound models from local JSON files:

```{r, eval = FALSE}
compound("path/to/Alfentanil-Model.json")
```


#### Explore Snapshots

Once you have imported a compound snapshot, you can explore its properties, such as compound properties, formulations, protocols, predefined simulations and all other information contained in the snapshot. For example, you can explore the formulations of Rifampicin using the following command:

```{r}
rifampicin$protocols[[1]]
```

This command outputs the formulations that are defined for the compound and that can be used in simulations.



### DDI projects

#### Create a DDI project

After importing the necessary compounds, you can create a DDI project. Here is a basic example of creating a DDI simulation with a victim compound (e.g., Rifampicin) and a perpetrator compound (e.g., Midazolam):

Some additional options can be passed during ddi creation:

```{r, warning=FALSE}
myDDI <- create_ddi(
  victim = rifampicin, 
  midazolam,
  options = list(
    import_simulations = FALSE,
    create_ddi_simulation = TRUE
  )
)
```

All options can be found in the `?create_ddi` help page.  

By default, the `create_ddi` function will generate a generic and ready-to-run ddi simulation.


#### Run the DDI simulations

Simulations defined in snapshots can easily be executed and their results retrieved using the `run_ddi()` function:

```{r}
results <- run_ddi(myDDI)
```
This will return a named list of data frames containing the simulation results for each simulation in the DDI project. You can access the results for a specific simulation using its name as a list index:

```{r, eval = FALSE}
head(results$`Generic DDI simulation`)
```
```{r , echo = FALSE}
knitr::kable(head(results$`Generic DDI simulation`))
```


Results can also be exported as csv and pkml files with the `path` and `exportPKML` arguments:

```{r, eval = FALSE}
results <- run_ddi(myDDI, path = "output_directory/", exportPKML = TRUE)
```

#### Get PK Parameters

PK parameters can be computed directly from the simulation results using `run_pk_analysis()` function:

```{r}
pk_analysis <- run_pk_analysis(myDDI)
```


As well as results, the PK statistics can be accessed by simulation name:

```{r, eval = FALSE}
head(pk_analysis$`Generic DDI simulation`)
```
```{r, echo=FALSE}
knitr::kable(head(pk_analysis$`Generic DDI simulation`))
```


#### Plot DDI simulations results

DDI simulations can be generated using the `plot_ddi_results()` function:

```{r} 
plots <- plot_ddi_results(myDDI)
```

Similarly plots can be accessed and visualized by simulation name:

```{r plot-ddi-results, out.width = "85%", fig.align="center", fig.width=6, fig.asp=0.618, dpi = 300}
plots$`Generic DDI simulation`
```

#### Export the DDI project as a Snapshot

You can also export the entire DDI simulation as a snapshot file for future use or for importing into other platforms such as PK-Sim:

```{r, eval=FALSE}
# Export the DDI simulation to a JSON snapshot file
export_ddi(myDDI, "path/to/Rifampicin-Midazolam-DDI.json")
```

### Modify DDI projects

Once created, DDI projects can be modified in several ways:
- Building blocks can be added or modified,
- Simulations can be added, modified or removed.

#### Create and add building blocks

Currently, the {cts} package only supports the creation of the following building blocks:
- Administration Protocols
- Formulations

##### Administration Protocols
Simple and advanced administration protocols can be created and added to DDI projects.

```{r}
# Create a simple oral protocol
oral_protocol <- create_protocol(
  name = "oral 10 mg",
  type = "oral", 
  interval = "single",
  dose = 10,
  dose_unit = "mg",
    water_vol_per_body_weight = 3.5 # Custom water volume per body weight
)

# Create an intravenous protocol with custom infusion time
iv_protocol <- create_protocol(
  name = "iv 300 mg (0.5 h)",
  type = "iv",
  interval = "single", 
  dose = 300,
  dose_unit = "mg",
  infusion_time = 0.5,
  infusion_time_unit = "h"
)

# Create a multiple dose oral protocol
multiple_oral <- create_protocol(
  name = "oral 100 mg BID",
  type = "oral",
  interval = "12-12", # Twice daily dosing
  dose = 100,
  dose_unit = "mg",
  end_time = 48, # Total duration of 48 hours
  water_vol_per_body_weight = 4.0 # Custom water volume per body weight
)
```

Advanced protocols allow for more complex dosing regimens by combining multiple administrations 
into schemas that can be repeated. Here's an example of a loading dose followed by maintenance doses:


```{r}
# Create an advanced protocol for a loading dose followed by maintenance doses
advanced_protocol <- 
  create_advanced_protocol(name = "Loading + Maintenance Protocol")

# Add first schema for loading dose (single high dose)
advanced_protocol <- add_schema(
  advanced_protocol,
  schema_name = "Loading Phase",
  start_time = 0,    # Start immediately
  start_time_unit = "h",
  rep_nb = 1,        # Just once
  time_btw_rep = 24, # Not relevant for single rep but required
  time_btw_rep_unit = "h"
)

# Create the loading dose administration
loading_dose <- create_protocol(
  name = "Loading Dose",
  type = "oral",
  interval = "single", # Must be single for advanced protocols
  dose = 200,         # Higher initial dose
  dose_unit = "mg",
  water_vol_per_body_weight = 3.5
)

# Add loading dose to the first schema
advanced_protocol <- add_administration(
  advanced_protocol,
  schema_name = "Loading Phase",
  administration = loading_dose
)

# Add second schema for maintenance doses
advanced_protocol <- add_schema(
  advanced_protocol,
  schema_name = "Maintenance Phase",
  start_time = 24,    # Start 24h after beginning
  start_time_unit = "h",
  rep_nb = 5,         # Repeat 5 times
  time_btw_rep = 12,  # Every 12 hours
  time_btw_rep_unit = "h"
)

# Create the maintenance dose administration
maintenance_dose <- create_protocol(
  name = "Maintenance Dose",
  type = "oral",
  interval = "single",
  dose = 100,         # Lower maintenance dose
  dose_unit = "mg",
  water_vol_per_body_weight = 3.5
)

# Add maintenance dose to the second schema
advanced_protocol <- add_administration(
  advanced_protocol,
  schema_name = "Maintenance Phase",
  administration = maintenance_dose
)

# Print the complete advanced protocol structure
advanced_protocol
```

Once created, protocols can be added to a DDI project using the `add_protocol()` function.

```{r}
add_protocol(myDDI, advanced_protocol)
```

They can also be removed from a DDI project using the `remove_protocol()` function:

```{r}
remove_protocol(myDDI, "Loading + Maintenance Protocol")
```

##### Formulations

Formulations define how a drug is released into the system. The `{cts}` package supports several types of formulations:
- Dissolved
- Weibull
- Lint80
- Particle
- Table
- Zero Order
- First Order

Here are examples of different formulation types:

```r
# Create a simple dissolved formulation
dissolved_form <- create_formulation(
  name = "Simple Solution",
  type = "dissolved"  # Immediately available drug
)

# Create a Weibull tablet formulation
weibull_form <- create_formulation(
  name = "Tablet Formulation",
  type = "weibull",
  dissolution_time = 240,     # Time to achieve 50% dissolution
  dissolution_time_unit = "min",
  lag_time = 10,             # Delay before dissolution starts
  lag_time_unit = "min",
  dissolution_shape = 0.92,   # Shape parameter
  suspension = FALSE          # Use as tablet, not suspension
)

# Create a particle formulation with polydisperse distribution
particle_form <- create_formulation(
  name = "Particle Formulation",
  type = "particle",
  thickness = 30,            # Unstirred water layer thickness
  thickness_unit = "µm",
  distribution_type = "poly", # Polydisperse distribution
  particle_size_distribution = "normal",
  radius = 10,              # Mean particle radius
  radius_unit = "µm",
  radius_sd = 3,            # Standard deviation of radius
  radius_sd_unit = "µm",
  radius_min = 1,           # Minimum particle size
  radius_min_unit = "µm",
  radius_max = 19,          # Maximum particle size
  radius_max_unit = "µm",
  n_bins = 5               # Number of size bins
)

# Create a zero-order release formulation
zero_order_form <- create_formulation(
  name = "Controlled Release",
  type = "zero",
  end_time = 120,          # Duration of release
  end_time_unit = "min"
)

# Create a first-order release formulation
first_order_form <- create_formulation(
  name = "First Order Release",
  type = "first",
  thalf = 30,              # Half-life of drug release
  thalf_unit = "min"
)

# Create a table-based release profile
table_form <- create_formulation(
  name = "Custom Release Profile",
  type = "table",
  tableX = c(0, 1, 2, 4, 8, 12),     # Time points (hours)
  tableY = c(0, 0.2, 0.4, 0.7, 0.9, 1.0),  # Fraction released
  suspension = TRUE
)
```

Each formulation type has specific parameters that control how the drug is 
released, they can be found in the `?create_formulation` help page.

#### Add formulations to a compound
Formulations can be added or removed from a ddi project using the `add_formulation()` function:

```r
# Add formulation to a compound
add_formulation(myDDI, weibull_form)
```

They can also be removed from a DDI project using the `remove_formulation()` function:

```r
# Remove formulation from a compound
remove_formulation(myDDI, "Tablet Formulation")
```


### Create new simulations

The `{cts}` package provides flexible ways to create and customize simulations through the `create_simulation()` function. Here's how to create and configure simulations:

#### Basic Simulation Creation

```r
# Create a basic simulation with individual
basic_sim <- create_simulation(
  simulation_name = "Basic DDI Study",
  victim = "Rifampicin",
  perpetrators = "Midazolam",
  individual = "European Reference"  # Use individual OR population, not both
)

# Create a simulation with population instead
pop_sim <- create_simulation(
  simulation_name = "Population DDI Study",
  victim = "Rifampicin",
  perpetrators = "Midazolam",
  population = "European Population"
)
```

#### Customizing Simulations

Once created, simulations can be customized in several ways:

##### 1. Adding Compounds

```r
# Add a new compound to the simulation
add_compound(basic_sim, 
  compound = "New Compound",
  protocol = "oral 100 mg",  # Optional protocol
  formulation = list(       # Optional formulation mapping
    Key = "Formulation 1",
    Name = "Tablet"
  )
)
```

##### 2. Setting Protocols

```r
# Change protocol for an existing compound
set_compound_protocol(basic_sim,
  compound = "Rifampicin",
  protocol = "iv 300 mg (0.5 h)",
  formulation = list(
    Key = "Formulation 2",
    Name = "IV Solution"
  )
)
```

##### 3. Configuring Output Intervals

```r
# Set a single output interval (replaces existing)
set_output_interval(basic_sim,
  start_time = 0,
  end_time = 24,
  resolution = 10,  # points per unit
  unit = "h"
)

# Add additional output interval
add_output_interval(basic_sim,
  start_time = 24,
  end_time = 48,
  resolution = 5,
  unit = "h"
)
```

#### 4. Managing Interactions and Processes

```r
# Add specific interactions for a compound
add_interactions(basic_sim,
  compound = "Rifampicin",
  interactions = c("CYP3A4", "P-gp")
)

# Add specific processes for a compound
add_processes(basic_sim,
  compound = "Rifampicin",
  processes = c("Hepatic Clearance", "Renal Clearance")
)
```

##### 5. Configuring Outputs

```r
# Set specific outputs to track (replaces existing)
set_outputs(basic_sim, c(
  "Organism|PeripheralVenousBlood|Rifampicin|Plasma (Peripheral Venous Blood)",
  "Organism|PeripheralVenousBlood|Midazolam|Plasma (Peripheral Venous Blood)"
))

# Add additional outputs to track
add_outputs(basic_sim,
  "Organism|Liver|Rifampicin|Intracellular"
)
```

#### Adding Simulations to DDI Projects

Once configured, simulations can be added to DDI projects with optional settings:

```r
# Add simulation to DDI project
add_simulation(myDDI, basic_sim, 
  options = list(
    add_interactions = TRUE,  # Automatically add relevant interactions
    add_processes = TRUE     # Automatically add relevant processes
  )
)

# Remove simulation from DDI project
remove_simulation(myDDI, "Basic DDI Study")
```
