---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)
```

# cts

<!-- badges: start -->

<!-- badges: end -->

The `{cts}` (Contraceptives DDI Trial Simulation Platform) package provides a comprehensive R-based framework for designing and simulating drug-drug interactions (DDI) involving contraceptive drugs using physiologically based pharmacokinetic (PBPK) models. This package allows researchers to conduct *in silico* DDI simulations programmatically, enabling more automated and reproducible workflows. This is particularly useful for performing batch simulations, and integrating DDI modeling into larger R-based data analysis pipelines, thus enhancing flexibility and efficiency in model development and simulation.

#### Key Features

-   Import and Explore Compound Models: Retrieve compound models from the OSP model library, a URL, or a local file in JSON format.
-   Simulate DDIs: Design and run DDI simulations between compounds, specifying dosing protocols, individual characteristics, and other simulation parameters.
-   Analyze Results: --- .
-   Seamless Integration: Supports the export of simulation snapshots compatible with PK-Sim, enabling further exploration and analysis.


## Installation

You can install the development version of `cts` from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("pak")
pak::pak("esqLABS/cts")
```

## Example

Below are some examples to demonstrate how to use the package. These steps will guide you through importing compound models, exploring their properties, and simulating drug-drug interactions.

### List Available Compound Building Blocks

The `{cts}`package can interface with the Open Systems Pharmacology (OSP) model library to retrieve compound models. The `list_compounds()` function lists all available compounds in the OSP model library.

```{r example, attr.output='style="max-height: 100px;"'}
library(cts)

# List available compounds building blocks
list_compounds()
```

This command outputs a list of compound names available in the OSP model library, which you can select to create your DDI simulations.

### Import Compound Snapshots

The core functionality of the package involves importing compound models. You can import a compound model in various ways:


#### A. From OSP Qualified Building Blocks

You can directly import pre-qualified compounds from the OSP model library:

```{r, include=FALSE}
rifampicin <- compound("Rifampicin")
midazolam <- compound("Midazolam")
```


#### B. From URL

For compound models hosted online, use the URL to import the model:

```{r, eval = FALSE}
compound("https://raw.githubusercontent.com/Open-Systems-Pharmacology/Alfentanil-Model/v2.2/Alfentanil-Model.json")
```


#### C. From Local File

You can also import compound models from local JSON files:

```{r, eval = FALSE}
compound("path/to/Alfentanil-Model.json")
```


### Explore Compound

Once you have imported a compound, you can explore its properties, such as formulations, protocols, and parameters:

```{r, include=FALSE}
# Explore the formulations of Rifampicin
rifampicin$formulations
```

This command outputs the formulations that are defined for the compound and that can be used in simulations.


## Create and Parameterize a DDI Simulation

After importing the necessary compounds, you can create a DDI simulation. Here is a basic example of creating a DDI simulation with a victim compound (e.g., Rifampicin) and a perpetrator compound (e.g., Midazolam):

```{r, warning=FALSE}
# Create a DDI simulation
myDDI <- create_ddi(victim = rifampicin, perpetrator = midazolam)
```

Next, you can parameterize the DDI simulation by setting dosing protocols and individual characteristics. For example, you can set a dosing protocol for the victim compound:

```{r, eval=FALSE}
# Define dosing protocol
protocol_1 <- protocol(dose = 250, dose_unit = "mg", start = 0, end = 30, interval = 1, time_unit = "days")

# Apply the protocol to the compound in the DDI simulation
set_protocol(myDDI, compound = "Rifampicin", protocol = protocol_1)
```

Here, a dosing protocol is defined and applied to Rifampicin, specifying a daily dose of 250 mg for 30 days.


## Run the DDI Simulation and Export Results

Finally, you can run the DDI simulation and export the results for further analysis:

```{r, eval=FALSE}
# Run the DDI simulation and export results
run_ddi(myDDI, path = "path/to/output", 
        format = "csv",  # Export results as a CSV file
        pkml = TRUE,     # Export the simulation as a PK-Sim project file
        plots = TRUE,    # Generate simulation plots
        stats = TRUE     # Compute and include PK statistics
)
```

This command runs the DDI simulation, saves the results to the specified path, generates relevant plots, and computes PK statistics, making it easy to analyze the interaction outcomes.


## Export the DDI Simulation as a Snapshot

You can also export the entire DDI simulation as a snapshot file for future use or for importing into other platforms such as PK-Sim:

```{r, eval=FALSE}
# Export the DDI simulation to a JSON snapshot file
export_ddi(myDDI, "path/to/Rifampicin-Midazolam-DDI.json")
```


## Dependencies

The `{cts}` package relies on several components from the Open Systems Pharmacology Suite:

- **R version 4.4.0 or higher**.
- **External R packages**: `{ospsuite}` for interaction with the PK-Sim core and `{rSharp}` for R and C# integration. These packages are installed automatically when you install `{cts}`.

For more details on the installation and dependency requirements for `{ospsuite}`, please visit the `{ospsuiteR}` GitHub page [here](https://github.com/Open-Systems-Pharmacology/OSPSuite-R).

