---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)
library(cts)
rifampicin <- compound(testthat::test_path("data/Rifampicin-Model.json"))
midazolam <- compound(testthat::test_path("data/Midazolam-Model.json"))
```

# cts

<!-- badges: start -->

<!-- badges: end -->

The `{cts}` (Contraceptives DDI Trial Simulation Platform) package provides a comprehensive R-based framework for designing and simulating drug-drug interactions (DDI) involving contraceptive drugs using physiologically based pharmacokinetic (PBPK) models. This package allows researchers to conduct *in silico* DDI simulations programmatically, enabling more automated and reproducible workflows. This is particularly useful for performing batch simulations, and integrating DDI modeling into larger R-based data analysis pipelines, thus enhancing flexibility and efficiency in model development and simulation.

#### Key Features

-   Import and Explore Compound Models: Retrieve compound models snapshots from the OSP model library, a URL, or a local file in JSON format.
-   Simulate DDIs: Design and run DDI simulations between compounds, specifying dosing protocols, individual characteristics, and other simulation parameters.
-   Analyze Results: In addition to simulation results, this packages provides an immediate and first glance at the data by providing basic plots and tables.
-   Seamless Integration: Supports the export of simulation snapshots compatible with PK-Sim, enabling further configuration, exploration and analysis.


## Installation

You can install the development version of `cts` from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("pak")
pak::pak("esqLABS/cts")
```

## Example

Below are some examples to demonstrate how to use the package. These steps will guide you through importing compound models, exploring their properties, and simulating drug-drug interactions.

### List Available Compound Building Blocks

The `{cts}`package can interface with the Open Systems Pharmacology (OSP) model library to retrieve compound models. The `list_compounds()` function lists all available compounds in the OSP model library.

```{r example, eval = FALSE, attr.output='style="max-height: 100px;"'}
library(cts)

list_compounds()
```
```{r, echo = FALSE, max.print=10, message=FALSE, warning=FALSE}
# List available compounds building blocks
cat(cli::cli_fmt(list_compounds(), collapse = T))
```


This command outputs a list of compound names available in the OSP model library, which you can select to create your DDI simulations.

### Import Compound Snapshots

The core functionality of the package involves importing compound models from snapshot files. You can import a compound model in various ways:


#### From OSP Qualified Building Blocks

You can directly import pre-qualified compounds from the OSP model library:

```{r, eval=FALSE}
rifampicin <- compound("Rifampicin")
midazolam <- compound("Midazolam")
```


#### From URL

For compound models hosted online, use the URL to import the model:

```{r, eval = FALSE}
compound("https://raw.githubusercontent.com/Open-Systems-Pharmacology/Alfentanil-Model/v2.2/Alfentanil-Model.json")
```


#### From Local File

You can also import compound models from local JSON files:

```{r, eval = FALSE}
compound("path/to/Alfentanil-Model.json")
```


### Explore Compound

Once you have imported a compound snapshot, you can explore its properties, such as compound properties, formulations, protocols, predefined simulations and all other information contained in the snapshot. For example, you can explore the formulations of Rifampicin using the following command:

```{r}
rifampicin$formulations
```

This command outputs the formulations that are defined for the compound and that can be used in simulations.


## Add new building blocks

Only administration protocols can be altered/added/removed from snapshots

```{r}
my_protocol <- create_protocol("my_protocol",
                type = "oral",
                interval = "single",
                dose = 200,
                water_vol_per_body_weight = 3.5)
```

```{r}
add_protocol(rifampicin, my_protocol)
```

```{r}
purrr::map_chr(rifampicin$protocols, "name")
```


## Create, Parameterize and Run a DDI Simulation

After importing the necessary compounds, you can create a DDI simulation. Here is a basic example of creating a DDI simulation with a victim compound (e.g., Rifampicin) and a perpetrator compound (e.g., Midazolam):

Some additional options can be passed during ddi creation:

```{r, warning=FALSE}
myDDI <- create_ddi(
  victim = rifampicin, perpetrator = midazolam,
  options = list(
    import_simulations = FALSE,
    create_ddi_simulation = TRUE
  )
)
```

All options can be found in the `?create_ddi` help page.  

By default, the `create_ddi` function will generate a generic and ready-to-run ddi simulation.


## Run the snapshot simulations

```{r}
results <- run_ddi(myDDI)
```
```{r}
head(results$`Generic DDI simulation`)
```
Results can also be exported as csv and pkml files with the `path` and `exportPKML` arguments:

```{r, eval = FALSE}
results <- run_ddi(myDDI, path = "output_directory/", exportPKML = TRUE)
```

## Export the DDI Simulation as a Snapshot

You can also export the entire DDI simulation as a snapshot file for future use or for importing into other platforms such as PK-Sim:

```{r, eval=FALSE}
# Export the DDI simulation to a JSON snapshot file
export_ddi(myDDI, "path/to/Rifampicin-Midazolam-DDI.json")
```


## What's Next

It is also possible to define new simulations, then add or remove them from the DDI object.

```{r, eval = FALSE}
# Work in progress
add_simulation(myDDI,
  name = "New Simulation",
  model = "4Comp",
  perpetrator = "midazolam",
  individual = "Male",
  victim_formulation,
  victim_protocol,
  perpetrator_formulation,
  perpetrator_protocol,
  ...
)

remove_simulation(myDDI, "New Simulation")
```

## Run the DDI Simulation and Export Results

Finally, you can run the DDI simulations and export the results for further analysis:

```{r, eval=FALSE}
# Run the DDI simulation and export results
run_ddi(myDDI,
  path = "path/to/output",
  format = "csv", # Export results as a CSV file
  pkml = TRUE, # Export the simulation as a PK-Sim project file
  plots = TRUE, # Generate simulation plots
  stats = TRUE # Compute and include PK statistics
)
```

This command runs the DDI simulation, saves the results to the specified path, generates relevant plots, and computes PK statistics, making it easy to analyze the interaction outcomes.



## Dependencies

The `{cts}` package relies on several components from the Open Systems Pharmacology Suite:

- **R version 4.4.0 or higher**.
- **External R packages**: `{ospsuite}` for interaction with the PK-Sim core and `{rSharp}` for R and C# integration. These packages are installed automatically when you install `{cts}`.

For more details on the installation and dependency requirements for `{ospsuite}`, please visit the `{ospsuiteR}` GitHub page [here](https://github.com/Open-Systems-Pharmacology/OSPSuite-R).

