---
title: "Creating Building Blocks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating Building Blocks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/building-blocks-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)
```

# Creating Building Blocks

The `{cts}` package allows you to create and customize various building blocks for your drug-drug interaction (DDI) simulations. This vignette demonstrates how to create different types of building blocks and add them to DDI projects.

## Loading the Package

First, load the `cts` package:

```{r setup}
library(cts)
```

## Creating Administration Protocols

Administration protocols define how compounds are administered in a simulation. The `{cts}` package supports both simple and advanced administration protocols.

### Simple Administration Protocols

Simple protocols can be created using the `create_protocol()` function:

```{r simple-protocols}
# Create a simple oral protocol
oral_protocol <- create_protocol(
  name = "oral 10 mg",
  type = "oral",
  interval = "single",
  dose = 10,
  dose_unit = "mg",
  water_vol_per_body_weight = 3.5 # Custom water volume per body weight
)

# Create an intravenous protocol with custom infusion time
iv_protocol <- create_protocol(
  name = "iv 300 mg (0.5 h)",
  type = "iv",
  interval = "single",
  dose = 300,
  dose_unit = "mg",
  infusion_time = 0.5,
  infusion_time_unit = "h"
)

# Create a multiple dose oral protocol
multiple_oral <- create_protocol(
  name = "oral 100 mg BID",
  type = "oral",
  interval = "12-12", # Twice daily dosing
  dose = 100,
  dose_unit = "mg",
  end_time = 48, # Total duration of 48 hours
  water_vol_per_body_weight = 4.0 # Custom water volume per body weight
)
```

### Advanced Administration Protocols

Advanced protocols allow for more complex dosing regimens by combining multiple administrations into schemas that can be repeated:

```{r advanced-protocols}
# Create an advanced protocol for a loading dose followed by maintenance doses
advanced_protocol <-
  create_advanced_protocol(name = "Loading + Maintenance Protocol")

# Add first schema for loading dose (single high dose)
advanced_protocol <- add_schema(
  advanced_protocol,
  schema_name = "Loading Phase",
  start_time = 0, # Start immediately
  start_time_unit = "h",
  rep_nb = 1, # Just once
  time_btw_rep = 24, # Not relevant for single rep but required
  time_btw_rep_unit = "h"
)

# Create the loading dose administration
loading_dose <- create_protocol(
  name = "Loading Dose",
  type = "oral",
  interval = "single", # Must be single for advanced protocols
  dose = 200, # Higher initial dose
  dose_unit = "mg",
  water_vol_per_body_weight = 3.5
)

# Add loading dose to the first schema
advanced_protocol <- add_administration(
  advanced_protocol,
  schema_name = "Loading Phase",
  administration = loading_dose
)

# Add second schema for maintenance doses
advanced_protocol <- add_schema(
  advanced_protocol,
  schema_name = "Maintenance Phase",
  start_time = 24, # Start 24h after beginning
  start_time_unit = "h",
  rep_nb = 5, # Repeat 5 times
  time_btw_rep = 12, # Every 12 hours
  time_btw_rep_unit = "h"
)

# Create the maintenance dose administration
maintenance_dose <- create_protocol(
  name = "Maintenance Dose",
  type = "oral",
  interval = "single",
  dose = 100, # Lower maintenance dose
  dose_unit = "mg",
  water_vol_per_body_weight = 3.5
)

# Add maintenance dose to the second schema
advanced_protocol <- add_administration(
  advanced_protocol,
  schema_name = "Maintenance Phase",
  administration = maintenance_dose
)

# Print the complete advanced protocol structure
advanced_protocol
```

## Creating Formulations

Formulations define how a drug is released into the system. The `{cts}` package supports several types of formulations:

```{r formulations}
# Create a simple dissolved formulation
dissolved_form <- create_formulation(
  name = "Simple Solution",
  type = "dissolved" # Immediately available drug
)

# Create a Weibull tablet formulation
weibull_form <- create_formulation(
  name = "Tablet Formulation",
  type = "weibull",
  dissolution_time = 240, # Time to achieve 50% dissolution
  dissolution_time_unit = "min",
  lag_time = 10, # Delay before dissolution starts
  lag_time_unit = "min",
  dissolution_shape = 0.92, # Shape parameter
  suspension = FALSE # Use as tablet, not suspension
)

# Create a particle formulation with polydisperse distribution
particle_form <- create_formulation(
  name = "Particle Formulation",
  type = "particle",
  thickness = 30, # Unstirred water layer thickness
  thickness_unit = "µm",
  distribution_type = "poly", # Polydisperse distribution
  particle_size_distribution = "normal",
  radius = 10, # Mean particle radius
  radius_unit = "µm",
  radius_sd = 3, # Standard deviation of radius
  radius_sd_unit = "µm",
  radius_min = 1, # Minimum particle size
  radius_min_unit = "µm",
  radius_max = 19, # Maximum particle size
  radius_max_unit = "µm",
  n_bins = 5 # Number of size bins
)

# Create a zero-order release formulation
zero_order_form <- create_formulation(
  name = "Controlled Release",
  type = "zero",
  end_time = 120, # Duration of release
  end_time_unit = "min"
)

# Create a first-order release formulation
first_order_form <- create_formulation(
  name = "First Order Release",
  type = "first",
  thalf = 30, # Half-life of drug release
  thalf_unit = "min"
)

# Create a table-based release profile
table_form <- create_formulation(
  name = "Custom Release Profile",
  type = "table",
  tableX = c(0, 1, 2, 4, 8, 12), # Time points (hours)
  tableY = c(0, 0.2, 0.4, 0.7, 0.9, 1.0), # Fraction released
  suspension = TRUE
)
```

Each formulation type has specific parameters that control how the drug is released. For more details, see the documentation with `?create_formulation`.

## Adding Building Blocks to DDI Projects

After creating building blocks, you can add them to your DDI projects.

### Setting Up a DDI Project

First, let's create a simple DDI project:

```{r setup-ddi, eval=FALSE}
# Import compound models
rifampicin <- compound("Rifampicin")
midazolam <- compound("Midazolam")

# Create a basic DDI project
myDDI <- create_ddi(victim = rifampicin, midazolam)
```

### Adding Protocols to a DDI Project

You can add protocols to a DDI project using the `add_protocol()` function:

```{r add-protocol, eval=FALSE}
# Add a simple protocol to the DDI project
add_protocol(myDDI, oral_protocol)

# Add an advanced protocol to the DDI project
add_protocol(myDDI, advanced_protocol)
```

### Removing Protocols from a DDI Project

You can remove protocols from a DDI project using the `remove_protocol()` function:

```{r remove-protocol, eval=FALSE}
# Remove a protocol from the DDI project
remove_protocol(myDDI, "oral 10 mg")
```

### Adding Formulations to a DDI Project

You can add formulations to a DDI project using the `add_formulation()` function:

```{r add-formulation, eval=FALSE}
# Add a formulation to the DDI project
add_formulation(myDDI, weibull_form)
```

### Removing Formulations from a DDI Project

You can remove formulations from a DDI project using the `remove_formulation()` function:

```{r remove-formulation, eval=FALSE}
# Remove a formulation from the DDI project
remove_formulation(myDDI, "Tablet Formulation")
```
