---
title: "Creating DDI Projects"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating DDI Projects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/ddi-projects-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)
```

# Creating Drug-Drug Interaction (DDI) Projects

The `{cts}` package provides a comprehensive framework for creating, running, and analyzing drug-drug interaction (DDI) simulations. This vignette demonstrates how to create DDI projects, run simulations, and analyze the results.

## Loading the Package

First, load the `cts` package:

```{r setup}
library(cts)
```

## Creating a DDI Project

The core function for creating a DDI project is `create_ddi()`. This function requires at least two compound models: a victim compound and one or more perpetrator compounds.

### Basic DDI Project Creation

Here's a simple example of creating a DDI project with one victim and one perpetrator:

```{r basic-ddi, eval=FALSE}
# Import compound models
rifampicin <- compound("Rifampicin")
midazolam <- compound("Midazolam")

# Create a basic DDI project
myDDI <- create_ddi(victim = rifampicin, midazolam)
```

By default, `create_ddi()` will generate a generic and ready-to-run DDI simulation.

### Customizing DDI Project Creation

You can customize the DDI project creation by providing options:

```{r custom-ddi, eval=FALSE}
# Create a DDI project with custom options
myDDI <- create_ddi(
  victim = rifampicin,
  midazolam,
  options = list(
    import_simulations = FALSE, # Don't import simulations from compound models
    create_ddi_simulation = TRUE # Create a generic DDI simulation
  )
)
```

Available options include:

- `import_simulations`: Whether to import simulations from the victim and perpetrator compound models (default: `FALSE`)
- `create_ddi_simulation`: Whether to create a generic simulation template (default: `TRUE`)

### Creating a DDI Project with Multiple Perpetrators

You can include multiple perpetrator compounds in a DDI project:

```{r multi-perpetrator, eval=FALSE}
# Import another compound model
ketoconazole <- compound("Ketoconazole")

# Create a DDI project with multiple perpetrators
multiDDI <- create_ddi(
  victim = midazolam,
  rifampicin,
  ketoconazole
)
```

## Running DDI Simulations

Once you've created a DDI project, you can run the simulations using the `run_ddi()` function:

```{r run-ddi, eval=FALSE}
# Run the DDI simulations
results <- run_ddi(myDDI)
```

This will return a named list of data frames containing the simulation results for each simulation in the DDI project. You can access the results for a specific simulation using its name as a list index:

```{r access-results, eval=FALSE}
# Access results for a specific simulation
head(results$`Generic DDI simulation`)
```

### Exporting Simulation Results

You can export the simulation results to files by providing a path:

```{r export-results, eval=FALSE}
# Run simulations and export results
results <- run_ddi(myDDI, path = "output_directory/", exportPKML = TRUE)
```

The `exportPKML` parameter controls whether PKML files are also exported.

## Analyzing DDI Simulation Results

### Calculating PK Parameters

You can calculate pharmacokinetic (PK) parameters from the simulation results using the `run_pk_analysis()` function:

```{r pk-analysis, eval=FALSE}
# Calculate PK parameters
pk_analysis <- run_pk_analysis(myDDI)

# Access PK parameters for a specific simulation
head(pk_analysis$`Generic DDI simulation`)
```

You can also export the PK analysis results to files:

```{r export-pk, eval=FALSE}
# Calculate and export PK parameters
pk_analysis <- run_pk_analysis(myDDI, path = "pk_results/")
```

### Plotting DDI Simulation Results

The `plot_ddi_results()` function generates plots of the simulation results:

```{r plot-ddi, eval=FALSE}
# Generate plots
plots <- plot_ddi_results(myDDI)

# Access a specific plot
plots$`Generic DDI simulation`
```

You can customize the plots by passing additional parameters:

```{r custom-plot, eval=FALSE}
# Generate plots with custom aggregation method
plots <- plot_ddi_results(myDDI, aggregationMethod = "median")
```

You can also generate plots for specific simulations:

```{r specific-plots, eval=FALSE}
# Generate plots for specific simulations
plots <- plot_ddi_results(myDDI, simulationNames = c("Generic DDI simulation"))
```

## Importing and Exporting DDI Projects

### Importing a DDI Project

You can import a DDI project from a JSON file:

```{r import-ddi, eval=FALSE}
# Import a DDI project
importedDDI <- import_ddi("path/to/DDI-Project.json")
```

### Exporting a DDI Project

You can export a DDI project to a JSON file for future use or for importing into other platforms such as PK-Sim:

```{r export-ddi, eval=FALSE}
# Export a DDI project
export_ddi(myDDI, "path/to/Rifampicin-Midazolam-DDI.json")
```
