---
title: "Creating and Managing Simulations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating and Managing Simulations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/simulations-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)
```

# Creating and Managing Simulations

The `{cts}` package provides flexible ways to create, customize, and manage simulations for drug-drug interaction (DDI) studies. This vignette demonstrates how to create simulations, customize their properties, and add them to DDI projects.

## Loading the Package

First, load the `cts` package:

```{r setup}
library(cts)
```

## Creating Basic Simulations

The `create_simulation()` function is the main entry point for creating simulations. At a minimum, you need to specify a simulation name, a victim compound, one or more perpetrator compounds, and either an individual or a population.

### Creating a Simulation with an Individual

```{r basic-individual}
# Create a basic simulation with an individual
basic_sim <- create_simulation(
  simulation_name = "Basic Individual DDI Study",
  victim = "Rifampicin",
  perpetrators = "Midazolam",
  individual = "European Reference"  # Use individual OR population, not both
)

# Print the simulation to see its structure
basic_sim
```

### Creating a Simulation with a Population

```{r basic-population}
# Create a simulation with a population instead
pop_sim <- create_simulation(
  simulation_name = "Population DDI Study",
  victim = "Rifampicin",
  perpetrators = "Midazolam",
  population = "European Population"
)

# Print the simulation to see its structure
pop_sim
```

### Creating a Simulation with Multiple Perpetrators

You can include multiple perpetrator compounds in a simulation:

```{r multi-perpetrator}
# Create a simulation with multiple perpetrators
multi_perp_sim <- create_simulation(
  simulation_name = "Multi-Perpetrator DDI Study",
  victim = "Rifampicin",
  perpetrators = c("Midazolam", "Ketoconazole"),
  individual = "European Reference"
)

# Print the simulation to see its structure
multi_perp_sim
```

## Customizing Simulations

Once you've created a basic simulation, you can customize it in various ways.

### Adding Compounds

You can add additional compounds to an existing simulation:

```{r add-compound}
# Add a new compound to the simulation
basic_sim <- add_compound(
  basic_sim, 
  compound = "Ketoconazole",
  protocol = "oral 100 mg",  # Optional protocol
  formulation = list(       # Optional formulation mapping
    Key = "Formulation",
    Name = "Tablet"
  )
)

# Print the updated simulation
basic_sim
```

### Setting Protocols for Compounds

You can set or change the protocol for a compound in the simulation:

```{r set-protocol}
# Change protocol for an existing compound
basic_sim <- set_compound_protocol(
  basic_sim,
  compound = "Rifampicin",
  protocol = "iv 300 mg (0.5 h)",
  formulation = list(
    Key = "Formulation",
    Name = "IV Solution"
  )
)

# Print the updated simulation
basic_sim
```

### Configuring Output Intervals

You can configure the time intervals for which simulation results are generated:

```{r output-intervals}
# Set a single output interval (replaces existing)
basic_sim <- set_output_interval(
  basic_sim,
  start_time = 0,
  end_time = 24,
  resolution = 10,  # points per unit
  unit = "h"
)

# Add an additional output interval
basic_sim <- add_output_interval(
  basic_sim,
  start_time = 24,
  end_time = 48,
  resolution = 5,
  unit = "h"
)

# Print the updated simulation
basic_sim
```

### Managing Interactions and Processes

You can specify which interactions and processes to include in the simulation:

```{r interactions-processes}
# Add specific interactions for a compound
basic_sim <- add_interactions(
  basic_sim,
  compound = "Rifampicin",
  interactions = c("CYP3A4", "P-gp")
)

# Add specific processes for a compound
basic_sim <- add_processes(
  basic_sim,
  compound = "Rifampicin",
  processes = c("Hepatic Clearance", "Renal Clearance")
)

# Print the updated simulation
basic_sim
```

### Configuring Outputs

You can specify which model outputs to track during the simulation:

```{r outputs}
# Set specific outputs to track (replaces existing)
basic_sim <- set_outputs(
  basic_sim, 
  c(
    "Organism|PeripheralVenousBlood|Rifampicin|Plasma (Peripheral Venous Blood)",
    "Organism|PeripheralVenousBlood|Midazolam|Plasma (Peripheral Venous Blood)"
  )
)

# Add additional outputs to track
basic_sim <- add_outputs(
  basic_sim,
  "Organism|Liver|Rifampicin|Intracellular"
)

# Print the updated simulation
basic_sim
```

## Adding Simulations to DDI Projects

Once you've created and customized a simulation, you can add it to a DDI project.

### Setting Up a DDI Project

First, let's create a simple DDI project:

```{r setup-ddi, eval=FALSE}
# Import compound models
rifampicin <- compound("Rifampicin")
midazolam <- compound("Midazolam")

# Create a basic DDI project
myDDI <- create_ddi(victim = rifampicin, midazolam)
```

### Adding a Simulation to a DDI Project

You can add a simulation to a DDI project using the `add_simulation()` function:

```{r add-simulation, eval=FALSE}
# Add simulation to DDI project
myDDI <- add_simulation(
  myDDI, 
  basic_sim, 
  options = list(
    add_interactions = TRUE,  # Automatically add relevant interactions
    add_processes = TRUE     # Automatically add relevant processes
  )
)

# Print the DDI project to see the added simulation
myDDI
```

### Removing a Simulation from a DDI Project

You can remove a simulation from a DDI project using the `remove_simulation()` function:

```{r remove-simulation, eval=FALSE}
# Remove simulation from DDI project
myDDI <- remove_simulation(myDDI, "Basic Individual DDI Study")

# Print the DDI project to confirm removal
myDDI
```