---
title: "Working with Compound Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with Compound Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/working-with-compounds",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)
```

# Working with Compound Models

The `{cts}` package provides comprehensive methods to import compound models for use in drug-drug interaction (DDI) simulations. This vignette demonstrates all available approaches to import compound models from different sources and explore their properties.

## Loading the Package

First, load the `cts` package:

```{r setup}
library(cts)
```

## Available Compound Building Blocks

The `{cts}` package can interface with the Open Systems Pharmacology (OSP) model library to retrieve compound models. You can list all available compounds in the OSP model library using the `list_compounds()` function:

```{r list-compounds}
# List available compounds building blocks
list_compounds()
```

This command outputs a list of compound names available in the OSP model library, which you can select to create your DDI simulations.

## Importing Compound Models

The `compound()` function is the main entry point for importing compound models. It supports multiple sources and formats, providing flexibility in how you obtain your models.

### From OSP Qualified Building Blocks

You can directly import pre-qualified compounds from the OSP model library by specifying the compound name:

```{r import-from-library, eval=FALSE}
# Import Rifampicin from the OSP model library
rifampicin <- compound("Rifampicin")

# Import Midazolam from the OSP model library
midazolam <- compound("Midazolam")
```

### From Local Files

You can import compound models from local JSON files:

```{r import-from-file, eval=FALSE}
# Import a compound model from a local file
local_compound <- compound("path/to/Rifampicin-Model.json")
```

### From URLs

You can also import compound models directly from URLs:

```{r import-from-url, eval=FALSE}
# Import a compound model from a URL
url_compound <- compound("https://raw.githubusercontent.com/Open-Systems-Pharmacology/Alfentanil-Model/v2.2/Alfentanil-Model.json")
```

## Exploring Compound Properties

After importing a compound model, you can explore its properties and structure.

### Basic Information

You can print a compound object to see its basic structure:

```{r print-compound, eval=FALSE}
# Print basic information about the compound
rifampicin
```

### Extracting Interactions

The `extract_interactions()` function allows you to identify all drug interactions defined in a compound model:

```{r extract-interactions, eval=FALSE}
# Extract all interactions for a compound
extract_interactions(rifampicin)

# Extract interactions for specific compounds
extract_interactions(rifampicin, compounds = "Rifampicin")

# Extract interactions without printing to console
interactions <- extract_interactions(rifampicin, quietly = TRUE)
```

### Extracting Processes

The `extract_processes()` function helps you identify all processes defined in a compound model:

```{r extract-processes, eval=FALSE}
# Extract all processes for a compound
extract_processes(rifampicin)

# Extract processes for specific compounds
extract_processes(rifampicin, compounds = "Rifampicin")

# Extract processes without printing to console
processes <- extract_processes(rifampicin, quietly = TRUE)
```

## Updating Compound Models

If you need to update a compound model to the latest version, you can use the `update_snapshots()` function:

```{r update-snapshots, eval=FALSE}
# Update a compound model to the latest version
updated_rifampicin <- update_snapshots(rifampicin)
```

## Exporting Compound Models

You can export a compound model to a JSON file using the `export()` method:

```{r export-compound, eval=FALSE}
# Export a compound model to a JSON file
rifampicin$export("path/to/exported-rifampicin.json")
```


