---
title: "CTS Package Technical Demo"
subtitle: "v1.1.0 - Drug-Drug Interaction (DDI) Simulation Platform"
date: today
format: 
  html:
    code-fold: false
    toc: true
    toc-depth: 3
    toc-float: true
    theme: cosmo
    highlight-style: github
    df-print: paged
    self-contained: true
knitr:
  opts_chunk: 
    results: hold
execute:
  echo: true
  warning: false
---

This document demonstrates the key features and workflow of the CTS package for simulating Drug-Drug Interactions (DDI). The demo covers loading compounds, creating DDI projects, defining protocols, and running simulations.

## Setup

First, we'll load the required packages. The CTS package is loaded using `devtools::load_all()` since we're working with a development version.

```{r}
#| label: setup
# Additional packages for this document
library(here)
library(knitr)

# Load the CTS package
devtools::load_all(here())
```

## 1. Available Compounds

The CTS package provides access to a library of pre-defined compound models from the Open Systems Pharmacology (OSP) suite. Let's explore what compounds are available using the `list_compounds()` function.

```{r}
#| label: compounds

# List available compounds
available_compounds <- list_compounds()
```

```{r, echo=FALSE}
#| tbl-cap: "Available Compounds in OSP Library"
# Display as formatted table
data.frame(
  Compound = available_compounds
)
```

## 2. Importing Compounds

For our DDI simulation, we'll import two compounds: Levonorgestrel (victim drug) and Itraconazole (perpetrator drug). The `compound()` function loads the model files and creates compound objects that can be used in simulations.

```{r}
#| label: import
#| class-output: txt

# Import example compounds
lng <- compound(here("tests/testthat/data/Levonorgestrel-Model.json"))
itr <- compound(here("tests/testthat/data/Itraconazole-Model.json"))

# Show compound information
itr
```

## 3. Creating DDI Project

A DDI project is created using `create_ddi()`, which sets up the framework for studying interactions between the victim and perpetrator drugs. The function automatically configures the simulation environment and can optionally create a default DDI simulation.

```{r}
#| label: ddi-project
#| class-output: txt

# Create DDI project
ddi_project <- create_ddi(
  victim = lng,
  itr, # Several compounds can be passed as separate arguments
  options = list(
    create_ddi_simulation = FALSE
  )
)

# Display project information
ddi_project
```

## 4. Formulations

### Existing Formulations
The DDI project includes default formulations for each compound. These formulations define how the drugs are administered and their release characteristics.

```{r}
#| label: existing formulations
#| class-output: txt

# List existing formulations
ddi_project$formulations
```

### New Formulations

We can create custom formulations using `create_formulation()`. Here, we'll create a Lint80 formulation with specific dissolution and lag time parameters.

```{r}
#| label: new formulation

# Create a new Lint80 formulation
lint8030 <- create_formulation(
  name = "Lint80-30",
  type = "lint80",
  dissolution_time = 30,
  dissolution_time_unit = "min",
  lag_time = 10,
  lag_time_unit = "min",
  suspension = FALSE
)
```

```{r}
#| label: add new formulation to the project
#| class-output: txt

# Add the new formulation to the DDI project using the add_formulation method
ddi_project$add_formulation(lint8030)

# List all formulations in the DDI project
ddi_project$formulations
```

## 5. Protocols

### Existing Protocols

Protocols define the dosing regimen for each compound. The DDI project comes with several pre-defined protocols that we can examine.

```{r}
#| label: existing protocols
#| class-output: txt

# List existing protocols
ddi_project$protocols[1:3]
```

### New Protocols

We can create both simple and advanced protocols. Simple protocols are created using `create_protocol()` for basic dosing regimens.

```{r}
#| label: new single dose protocol
#| class-output: txt

# Create a new dose protocol for daily administration over 30 days
oral_150ug_over_30d <- create_protocol(
  name = "oral_150ug_over_30d",
  type = "oral",
  interval = "24",
  dose = 0.15,
  dose_unit = "mg",
  end_time = 30,
  end_time_unit = "day(s)",
  water_vol_per_body_weight = 3.5
)

oral_150ug_over_30d
```

For more complex dosing regimens, we use `create_advanced_protocol()` which allows for multiple dosing phases and schedules.

```{r}
#| label: new advanced protocol
#| class-output: txt
#| results: hold

# Create an advanced protocol with a 14-day lead-in phase
oral_750u_start_14d <- create_advanced_protocol("oral_750u_start_14d") |>
  # Schema definition for the DDI phase
  add_schema(
    schema_name = "DDI phase",
    start_time = 14, # Start after 14 days
    start_time_unit = "day(s)",
    rep_nb = 16, # 16 days duration
    time_btw_rep = 24, # Once daily
    time_btw_rep_unit = "h"
  ) |>
  # Dose definition
  add_administration(
    schema_name = "DDI phase",
    create_protocol(
      name = "DDI Dose",
      type = "oral",
      interval = "single",
      dose = 0.750, # Maintenance dose
      dose_unit = "mg",
      water_vol_per_body_weight = 3.5
    )
  )

oral_750u_start_14d
```

### Add Protocols to DDI Project

The created protocols are added to the DDI project for use in simulations.

```{r}
#| label: add protocols to ddi project
#| class-output: txt

# Add both protocols to the DDI project
ddi_project$add_protocol(oral_150ug_over_30d)
ddi_project$add_protocol(oral_750u_start_14d)

ddi_project$protocols[13:14]
```

## 6. Create a new simulation

Now we'll create a custom simulation using `create_simulation()`. This defines the basic parameters of our DDI study including the compounds involved and the study population.

```{r}
#| label: create a new simulation
#| class-output: txt

new_simulation <- create_simulation(
  simulation_name = "LNG_ITR_DDI_150_750_30d_14d",
  victim = "Levonorgestrel 1",
  perpetrators = "Itraconazole",
  individual = "Woman"
)

new_simulation
```

## 7. Customise Simulation

### Add compound protocols

We need to specify which protocols and formulations to use for each compound in the simulation. This links the dosing regimens we created earlier to specific compounds.

```{r}
#| label: add compound protocols
#| class-output: txt

new_simulation |>
  set_compound_protocol(
    compound = "Levonorgestrel 1",
    protocol = "oral_150ug_over_30d",
    formulation = "Lint80-30"
  ) |>
  set_compound_protocol(
    compound = "Itraconazole",
    protocol = "oral_750u_start_14d",
    formulation = "IR Dissolved"
  )

new_simulation
```

::: {.callout-note}
`set_compound_protocol()` is being reworked to be easier to use. The current version requires specifying the formulation as a list of key-value pairs. The new version will allow direct assignment using the formulation name.
:::

### Set output Interval

The output interval defines when and how frequently the simulation calculates results. We can set multiple intervals with different resolutions to capture both rapid changes and long-term effects.

```{r}
#| label: set output interval
#| class-output: txt


# Set initial output interval with higher resolution
new_simulation |>
  set_output_interval(
    start_time = 0,
    end_time = 14,
    resolution = 24 * 5, # 5 points per hour
    unit = "day(s)"
  )

# Add another interval with different resolution
new_simulation |>
  add_output_interval(
    start_time = 15,
    end_time = 30,
    resolution = 24 * 10, # 10 points per hour
    unit = "day(s)"
  )

new_simulation$output_schema
```

### Define simulated outputs

Here we specify which quantities to track during the simulation. For DDI studies, we typically monitor plasma concentrations of both compounds to observe their interaction.

```{r}
#| label: define outputs
#| class-output: txt

# Set outputs to track plasma concentrations
new_simulation <- set_outputs(
  new_simulation,
  c(
    "Organism|Venous Blood|Plasma|Levonorgestrel 1|Concentration",
    "Organism|Venous Blood|Plasma|Itraconazole|Concentration"
  )
)

# Display the simulation configuration
new_simulation$output_selections
```

## 8. Add Simulation to DDI Project

Now that we have configured our simulation, we can add it to our DDI project. This step registers the simulation settings and prepares it for execution.

```{r}
#| label: add simulation to ddi project
#| class-output: txt

# Add the simulation to the DDI project
ddi_project <- add_simulation(ddi_project, new_simulation)

# Verify the simulation was added
ddi_project
```

## 9. Run Simulation

Finally, we can run the simulation using the `run()` method. This will execute the DDI study with all our configured parameters and return the results.

```{r}
#| label: run simulation
#| class-output: txt
#| eval: false

# Run the simulation
results <- ddi_project$run()

# Display basic information about the results
results

# Plot the plasma concentrations over time
plot(results)
```

